### 2015-09-16 poster/poster.Rnw
clarify figure legends and make other changes Monica suggested

\documentclass[a0paper,landscape]{tikzposter}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{booktabs}
% \usepackage{multirow}
\usepackage{acronym}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
\addbibresource{references.bib}
% \usepackage[nomargin,inline,draft]{fixme}
% \newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
% \renewcommand*{\bibfont}{\small}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.  Use showboxes option to show outlines of textboxes.
% \usepackage[absolute]{textpos}
\usepackage[absolute,showboxes]{textpos}

\usepackage{multirow}
\usepackage{array}

\title{Identifying the Tissue of Origin of Extracellular Vesicles Using RNA Expression Signatures}
\author{Don L. Armstrong, Monica Uddin, Derek Wildman}
\institute{Institute for Genomic Biology, Computing Genomes for Reproductive Health, University of Illinois, Urbana-Champaign}
\usetheme{Default}

% use this
% http://www.identitystandards.illinois.edu/graphicstandardsmanual/othermedia/posterpresentation.html
\definecolorstyle{UofI}{
  \definecolor{boldBlue}{HTML}{002058}
  \definecolor{boldOrange}{HTML}{E87722}
  \definecolor{professionalBlue}{HTML}{606EB2}
  \definecolor{professionalOrange}{HTML}{D45D00}
  }{
    \colorlet{backgroundcolor}{professionalBlue}
    \colorlet{framecolor}{black}
    \colorlet{titlefgcolor}{boldBlue}
    \colorlet{titlebgcolor}{white}
    \colorlet{blockbodybgcolor}{white}
    \colorlet{blocktitlebgcolor}{professionalOrange}
    \colorlet{blocktitlefgcolor}{boldBlue}
    }
\usecolorstyle{UofI}
\defineblockstyle{UofIMinimal}{
    titlewidthscale=1, bodywidthscale=1, titleleft,
    titleoffsetx=0pt, titleoffsety=0pt, bodyoffsetx=0pt, bodyoffsety=0pt,
    bodyverticalshift=0pt, roundedcorners=0, linewidth=0.2cm,
    titleinnersep=1cm, bodyinnersep=1cm
}{
    \begin{scope}[line width=\blocklinewidth, rounded corners=\blockroundedcorners]
       \ifBlockHasTitle %
           \draw[draw=none,fill=blockbodybgcolor]
               (blockbody.south west) rectangle (blocktitle.north east);
           \draw[color=blocktitlebgcolor, loosely dashed]
               (blocktitle.south west) -- (blocktitle.south east);%
       \else
             \draw[draw=none,fill=blockbodybgcolor]
                 (blockbody.south west) rectangle (blockbody.north east);
        \fi
    \end{scope}
}

\useblockstyle{UofIMinimal}
\definebackgroundstyle{UofIVertical}{
    \draw[line width=0pt, bottom color=backgroundcolor, 
    top color=white] (bottomleft) rectangle (topright);
}
\usebackgroundstyle{UofIVertical}
\definetitlestyle{UofITitle}{
    width=\paperwidth, roundedcorners=0, linewidth=0pt, innersep=12mm,
    titletotopverticalspace=5mm, titletoblockverticalspace=10mm,
    titlegraphictotitledistance=10pt
}{}
\usetitlestyle{UofITitle}

\makeatletter
\renewcommand\TP@maketitle{%
   \begin{minipage}{0.8\textwidth}
        \color{titlefgcolor}
        {\bfseries \Huge \sc \@title \par}
        \vspace*{1em}
        {\huge \@author \par}
        \vspace*{1em}
        {\LARGE \@institute}
    \end{minipage}%
    \hfill
    \begin{minipage}{0.2\linewidth}
      \includegraphics[width=0.8\linewidth,height=6cm,keepaspectratio]{figures/igb_wordmark}
    \end{minipage}
}
\makeatother

\begin{document}
\acrodef{EV}[EV]{Extracellular Vesicle}
\acrodef{SVM}{Support Vector Machine}
\acrodef{ROC}{Receiver Operating Characteristic}
\acrodef{KNN}{K Nearest Neighbors}
\acrodef{SRA}{Sequence Read Archive}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="0.20\\columnwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
#opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(digits=2)
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
library("grid")
library("xtable")
library("caret")
library("pROC")

capwords <- function(s, strict = FALSE) {
    cap <- function(s) paste(toupper(substring(s, 1, 1)),
                             {s <- substring(s, 2); if(strict) tolower(s) else s},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}
@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced")))>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
@

\maketitle

<<specific.markers.plot.function,echo=FALSE,results="hide",warning=FALSE>>=
specific.markers.plot <- function(tissue.name,genes) {
    setkey(interesting.gene.reads.wide,"gene_short_name")
    if (missing(genes))
        genes <- as.character(tissue.specific.genes[tissue==tissue.name,gene])
    interesting.tissues <- 
        colnames(interesting.gene.reads.wide)[
           which(apply(apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
                             function(x){order(-x)==2}),1,any))+2]
     interesting.tissues <- 
        c("gene_short_name",tissue.name,interesting.tissues)

    tissue.specific.gene.reads <- 
        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
                        id="gene_short_name",
                        variable.name="tissue",
                        value.name="FPKM"
                        ))
    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
    return(ggplot(tissue.specific.gene.reads,
                  aes(y=gene_short_name,x=tissue))
           + geom_tile(aes(fill=`log_2(FPKM)`))
           + scale_fill_gradient(limits=c(0,17),
                                 low="white",
                                 high="#002058",
                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
           + scale_x_discrete("", expand = c(0, 0))
           + scale_y_discrete("", expand = c(0, 0))
           + theme_grey(base_size=14)
           + theme(legend.position = "top",
               axis.ticks = element_blank(), 
               plot.margin=unit(c(1,1,1,1),"mm"),
               axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
           )
}
@

\begin{columns}

  \column{0.25}

  \block{Introduction}{\acp{EV} are small (40-1000nm) compartments
    that can be found in blood, saliva, and urine. \acp{EV} originate
    from the plasma membrane or multivesicular endosome of many cell
    types and contain proteins, lipids, mRNA, miRNA, and many other
    non-coding RNAs. Because \acp{EV} contain mRNA from the cell of
    origin and can be obtained without invasive procedures, they are a
    promising remote sensor of the transcriptome of tissues that
    would otherwise be inaccessible.

    One requirement for remotely sensing the transcriptome is
    identifying the source tissue of a specific \ac{EV}. To identify
    the source tissue of a specific \ac{EV}, we have identified
    tissue-specific miRNA and mRNA markers for multiple organs and
    tissues from existing publicly-available RNAseq transcriptomes by
    calculating a tissue specificity index.

    The tissue specificity index for gene $g$ ($\tau_g$) is

    \begin{equation}
      \tau_g = \frac{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}{N-1}
    \end{equation}
    \label{eq:tissue_specificity_index}

    where $N$ is the number of tissues, $X_{i,g}$ is the mean
    expression of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is
    the maximal expression of gene $g$ across all tissues.

    we we trained a multiclass \ac{SVM} to accurately identify the
    tissue of origin, and present the results of the identification of
    reads from uterine and placental origin here. Tables showing
    tissue specific genes from all %
    %
    \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
    % 
    tissues can be found on \url{http://dla2.us/p/em2015}.

%     Our approach identified MIR1914, MIR126, SNORD12, and SNORD5 as
%     uterus specific. CGB, LGALS13, and members of the PSG family were
%     identified as placenta specific. The accuracy of the classifier
%     was verified using RNAseq transcriptome data not in the training
%     set. We show the applicability of the classifier to simulated EVs
%     by down sampling reads to $10^5$ reads and biasing the reads to
%     account for preferential association of RNA motifs with EV
%     packaging machinery. When coupled with single-EV sequencing, our
%     classifier enables clinicians to monitor disease-specific markers
%     over the course of treatment without biopsies, increasing the
%     ability of clinicians to adapt treatment modalities to the
%     individual patient. 
  }

  \block{Methods}{ %
    \Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]} %
    %
    RNAseq reads from %
    % 
    \Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
    % 
    samples (supplemented with FPKMs from
    \Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
      nrow(predictions.caret.svm[["uterus"]][["testing"]]) -
      length(star.logs[grepl("Number of input reads",field),value])} %
    %
    additional samples without raw data in the \ac{SRA})
    % 
    corresponding to %
    \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
    different tissues were obtained from the Roadmap Epigenomics
    \cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping} and GTEx
    \cite{Gtexconsortium.2015:Thegenotypetissueexpression} projects,
    as well as uterine (GSE50599,
    \cite{Chan.ea2014:Assessmentofmyometrialtranscriptome}) and
    placenta (GSE66622,
    \cite{Hughes.ea2015:Evaluatingintra-andinter-individual}) samples
    from the \ac{SRA}.

    Reads were aligned to \textit{Homo sapiens} reference genome
    GRCh38 with Ensembl annotation release 80 using STAR v2.4.2a
    \cite{Dobin.ea2013:STARultrafastuniversalRNA-seq} and called using
    Tophat 2 v2.0.10
    \cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}. The
    average percentage of uniquely mapped reads was %
    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value])))}
    and the average total mapping percentage was %
    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value]))+as.numeric(gsub("%","",star.logs[field=="% of reads mapped to multiple loci",value])))}.
    % 
    For samples where the reads were not publicly available (GTEx and
    some Roadmap Epigenomics samples), the RPKM were taken from
    publicly available result files (see code for details).

    The \ac{SVM} was generated using caret
    \cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
    \cite{Karatzoglou2004:kernlab} with 10 cycles of 10 repeats and a
    tune length of 8 with a training set representing $\approx 75\%$ of the
    samples ($\approx 2450$ samples total, including $\approx 35$
    uterine or $\approx 60$ placenta) in each of the groups chosen at
    random, with the remaining samples making up a testing set
    ($\approx 820$ samples).

}

\column{0.25}
  \block{Uterus Specific Genes}{

{\centering
% <<uterus.specific.markers.table,results="asis",echo=FALSE>>=
% temp <- tissue.specific.genes[tissue=="uterus",list(gene,specificity,gi.max)]
% setorder(temp,-gi.max)
% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
% print(xtable(temp,
%              caption="Uterus specific genes",label="tab:uterus_specific_genes"),
%       floating=FALSE,
%       include.rownames=FALSE,
%       )
% rm(temp)
% @ 
% }

<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("uterus")
      + xlab("Uterus Specific Genes")
      + ylab("Tissue"))
@
}
\captionof{figure}{Expression of uterus specific genes in uterus and
  tissues with the second highest expression for those genes.
  Expression is $\log_2 \mathrm{FPKM}$ and genes are tissue specific
  if they have a tissue specific expression $\tau_g \ge 0.98$. The
  known uterus specific gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp)
  and AL356310.1 (94bp) are highly uterus specific.}

}

  \block{Placenta Specific Genes}{
{\centering
% <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
% temp <- tissue.specific.genes[tissue=="placenta" & gi.max >= 1000,
%                               list(gene,specificity,gi.max)]
% setorder(temp,-gi.max)
% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
% print(xtable(temp,
%              caption="Placenta specific genes",label="tab:placenta_specific_genes"),
%       floating=FALSE,
%       include.rownames=FALSE,
%       )
% rm(temp)
% @

<<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("placenta",
                            genes=as.character(tissue.specific.genes[tissue=="placenta" & 
                                                                         gi.max >= 1024
                               ,gene]))
      + xlab("Uterus Specific Genes")
      + ylab("Tissue"))
@

}
\captionof{figure}{Expression of placenta specific genes in placenta
  and tissues with the second highest expression for those genes.
  Expression is $\log_2 \mathrm{FPKM}$ and genes are tissue specific
  if they have a tissue specific expression $\tau_g \ge 0.98$ and
  $\mathrm{FPKM} \ge 1024$ in placenta. The known placenta specific
  PSG genes and CSH genes are all highly placenta specific, and short
  RNAs mir-126, mir-622, and PAPPA-AS1 are fairly placenta specific.}
}

\column{0.25}

\block{SVM to distinguish tissues}{
  % Describe SVM approach

  % Figure showing the samples

  % Ability to identify placenta

  % Ability to identify uterus

<<placenta.roc.curve,fig.width=8,fig.height=4,echo=FALSE>>=
layout(matrix(1:2,1,2))
pl.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"actual"])-1,
             as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"placenta"]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
pl.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"actual"])-1,
              as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"placenta"]),
              percent=TRUE,
              col="#E87722")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",
       legend=c("Placenta Testing Set", "Placenta Training Set"),
       col=c("#002058", "#E87722"),
       lwd=2)

ut.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"actual"])-1,
             as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"uterus"]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
ut.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"actual"])-1,
              as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"uterus"]),
              percent=TRUE,
              col="#E87722")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",
       legend=c("Uterus Testing Set", "Uterus Training Set"),
       col=c("#002058", "#E87722"),
       lwd=2)
@ 
\captionof{figure}{\ac{ROC} curve of the placenta \ac{SVM} (left) and
  uterus \ac{SVM} right) . Both \ac{SVM} are highly effective at
  distinguishing between samples of uterus origin and non-uterus
  origin.}
\label{fig:plac_ut_full_roc}
}

\block{Identification of tissue in reduced reads}{
  % Describe reduced read approach
  % 
<<placenta.roc.reduced.curve,fig.width=8,fig.height=10,echo=FALSE>>=
layout(matrix(1:2,2,1))
for (a in c("placenta","uterus")) {
pl.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,"actual"])-1,
             as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,a]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
pl.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,a]),
              percent=TRUE,
              col="#E87722")
pl.roc.obj3 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,a]),
              percent=TRUE,
              col="green")
pl.roc.obj4 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"placenta"]),
              percent=TRUE,
              col="brown")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
       col=c("brown","green","#002058", "#E87722"),
       lwd=2)
}
@ 
\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta and
  Uterus samples with the indicated number of total reads. Reads were
  reduced by sampling reads for each sample weighted by the FPKM
  measured for each sample, and run using the same trained SVM
  depicted in Figure \ref{fig:plac_ut_full_roc}. %
  % 
  The placenta \ac{SVM} is fairly accurate at distinguishing between
  samples of placental origin and non-placental origin even when the
  number of reads is very low (10). The uterus \ac{SVM} requires many
  more reads in order to achieve similar accuracy. This is likely due
  to the many fewer uterus specific genes.}
}

\column{0.25}

\block{Conclusions}{
  \begin{itemize}
  \item Pregnancy-specific glycoproteins (PSG1-8) are specific to
    placental tissues, and PAEP is specific to uterus, as previously
    described.
%   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
  \item \ac{SVM} are able to accurately identify the tissue of origin
    for uterus and placenta.
  \item Very few sequenced reads are required to accurately identify
    tissues of Placental origin. We are able to detect 100\% of the
    placenta tissues with 80\% specificity with just 10 reads.
  \item We can easily identify exosomes originating from the placenta,
    as even the smallest exosomes can contain 70 RNA molecules
    \cite{Li.ea2014:AnalysisofRNAcontent}.
  \item Non-maternal variants may also be able to distinguish between
    placental reads and maternal reads.
  \end{itemize}
}

\block{References}{

  \begin{center}
    \mbox{}\vspace{-\baselineskip}
  \printbibliography[heading=none]
\end{center}
}

\block{Code and Datasets}{
The code and underlying data for this poster can be found at
\url{http://dla2.us/p/em2015}

% 
}
\block{}{
  \includegraphics[width=\linewidth,height=6cm,keepaspectratio]{figures/uofi_mark}
}

\end{columns}

\end{document}