### 2015-09-25 slides/slides.Rnw
add start of slides for presentation based on the poster

\documentclass[ignorenonframetext]{beamer}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{booktabs}
% \usepackage{multirow}
\usepackage{acronym}
\usepackage{qrcode}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
\addbibresource{references.bib}
% \usepackage[nomargin,inline,draft]{fixme}
% \newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
% \usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\renewcommand*{\bibfont}{\tiny}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.  Use showboxes option to show outlines of textboxes.
% \usepackage[absolute]{textpos}
\usepackage[absolute,showboxes]{textpos}

\usepackage{multirow}
\usepackage{array}

\mode<presentation>{ 
  \usetheme{CambridgeUS}
  % http://identitystandards.illinois.edu/graphicstandardsmanual/generalguidelines/colors.html
  \definecolor{ilboldblue}{HTML}{002058}
  \definecolor{ilboldorange}{HTML}{E87722}
  \definecolor{ilblue}{HTML}{606EB2}
  \definecolor{ilorange}{HTML}{D45D00}
  \setbeamercolor{alerted text}{fg=ilboldorange}
  \setbeamercolor*{palette primary}{fg=ilblue,bg=ilorange}
  \setbeamercolor*{palette secondary}{fg=ilblue!20!white,bg=ilorange}
  \setbeamercolor*{palette tertiary}{bg=ilblue,fg=ilorange}
  \setbeamercolor*{palette quaternary}{fg=ilblue,bg=ilorange}

  \setbeamercolor*{sidebar}{fg=ilorange,bg=ilboldblue}

  \setbeamercolor*{palette sidebar primary}{fg=ilblue!10!white,bg=ilorange}
  \setbeamercolor*{palette sidebar secondary}{fg=ilorange}
  \setbeamercolor*{palette sidebar tertiary}{fg=ilblue}
  \setbeamercolor*{palette sidebar quaternary}{fg=ilorange}

  % \setbeamercolor*{titlelike}{parent=palette primary}
  \setbeamercolor{titlelike}{parent=palette primary,fg=ilboldblue,bg=ilorange}
  \setbeamercolor{frametitle}{fg=ilboldorange,bg=ilblue!80!white}
  \setbeamercolor{frametitle right}{fg=ilboldblue,bg=ilorange}

  \setbeamercolor*{separation line}{}
  \setbeamercolor*{fine separation line}{}
  \setbeamercovered{transparent}  
  \logo{\begin{tikzpicture}% Pale figure
      {\node[opacity=0.8]{\IfFileExists{figures/igb_wordmark.pdf}{\includegraphics[width=2cm,height=1cm,keepaspectratio]{figures/igb_wordmark}}{}%
        };}%
    \end{tikzpicture}}
}

\title{Identifying Extracellular Vesicles of Placenta and Uterine 
  Origin Using RNA Expression Signatures}
\author{Don L. Armstrong, Monica Uddin, Derek Wildman}
\institute{Institute for Genomic Biology, Computing Genomes 
  for Reproductive Health, University of Illinois, Urbana-Champaign}
\subject{RNA expression}

\begin{document}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.8\\textheight",out.extra="keepaspectratio")
#opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(digits=2)
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
library("grid")
library("xtable")
library("caret")
library("pROC")

capwords <- function(s, strict = FALSE) {
    cap <- function(s) paste(toupper(substring(s, 1, 1)),
                             {s <- substring(s, 2); if(strict) tolower(s) else s},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}
@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced")))>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
@

\IfFileExists{figures/relevant_xkcd.png}{\frame[plain]{\centering \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{figures/relevant_xkcd.png}

    \url{https://xkcd.com/1217/}}}

\frame[plain]{\titlepage}

<<specific.markers.plot.function,echo=FALSE,results="hide",warning=FALSE>>=
specific.markers.plot <- function(tissue.name,genes) {
    setkey(interesting.gene.reads.wide,"gene_short_name")
    if (missing(genes))
        genes <- as.character(tissue.specific.genes[tissue==tissue.name,gene])
    interesting.tissues <- 
        colnames(interesting.gene.reads.wide)[
         apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
                             function(x){order(-x)[2]+2})]

    interesting.tissues <- 
        c("gene_short_name",tissue.name,unique(interesting.tissues))

    tissue.specific.gene.reads <- 
        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
                        id="gene_short_name",
                        variable.name="tissue",
                        value.name="FPKM"
                        ))
    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
    gene.levels <- 
        unique(as.character(tissue.specific.genes[tissue==tissue.name][order(specificity),gene]))
    tissue.specific.gene.reads[,gene_short_name:=factor(gene_short_name,levels=gene.levels)]
    return(ggplot(tissue.specific.gene.reads,
                  aes(y=gene_short_name,
                          x=tissue))
           + geom_tile(aes(fill=`log_2(FPKM)`))
           + scale_fill_gradient(limits=c(0,17),
                                 low="white",
                                 high="#002058",
                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
           + scale_x_discrete("", expand = c(0, 0))
           + scale_y_discrete("", expand = c(0, 0))
           + theme_grey(base_size=14)
           + theme(legend.position = "top",
                   axis.ticks = element_blank(),
                   panel.grid=element_blank(),
                   panel.border=element_blank(),
                   panel.background=element_blank(),
                   plot.background=element_blank(),
                   plot.margin=unit(c(1,1,1,1),"mm"),
                   axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
           )
}
@

\section{Introduction}

\begin{frame}{Introduction: \acsp{EV}}
  \acp{EV} are
  \begin{itemize}
  \item small (40-1000nm) compartments
  \item found in blood, saliva, and urine. 
  \item originate from the plasma membrane or multivesicular endosome
  \item from many cell types
  \item contain proteins, lipids, mRNA, miRNA, and other ncRNAs.
  \item \alert{remote sensor of the transcriptome of inaccessible tissues}
  \end{itemize}
\end{frame}

\subsection{Identifying source tissue of \acp{EV}}

\begin{frame}{Identifying source tissue of \acp{EV}}

  One requirement for remotely sensing the transcriptome is
  identifying the source tissue of a specific \ac{EV}. To identify the
  source tissue of a specific \ac{EV}, we have identified miRNA and
  mRNA markers which are characteristic of multiple organs and tissues
  from existing publicly-available RNAseq transcriptomes by
  calculating a tissue specificity index.

  The tissue specificity index for gene $g$ ($\tau_g$) is

  \begin{equation}
    \tau_g = \frac{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}{N-1}
  \end{equation}
  \label{eq:tissue_specificity_index}

  where $N$ is the number of tissues, $X_{i,g}$ is the mean expression
  of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is the maximal
  expression of gene $g$ across all tissues.

 \end{frame}

 \begin{frame}{\ac{SVM}}
    We also trained a multiclass \ac{SVM} to accurately identify the
    tissue of origin, and present the results of the identification of
    reads from uterine and placental origin here. Tables showing
    tissue specific genes from all %
    %
    \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
    % 
    tissues can be found on \url{http://dla2.us/p/em2015}.

%     Our approach identified MIR1914, MIR126, SNORD12, and SNORD5 as
%     uterus specific. CGB, LGALS13, and members of the PSG family were
%     identified as placenta specific. The accuracy of the classifier
%     was verified using RNAseq transcriptome data not in the training
%     set. We show the applicability of the classifier to simulated EVs
%     by down sampling reads to $10^5$ reads and biasing the reads to
%     account for preferential association of RNA motifs with EV
%     packaging machinery. When coupled with single-EV sequencing, our
%     classifier enables clinicians to monitor disease-specific markers
%     over the course of treatment without biopsies, increasing the
%     ability of clinicians to adapt treatment modalities to the
%     individual patient. 
\end{frame}

\section{Methods}
\begin{frame}{Methods}
  \begin{itemize}
  % 
  \item \Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]} %
    %
    RNAseq reads (\Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]/length(star.logs[grepl("Number of input reads",field),value])} average per sample)%
    % 
  \item  \Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
    % 
    samples (supplemented with FPKMs from
    \Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
      nrow(predictions.caret.svm[["uterus"]][["testing"]]) -
      length(star.logs[grepl("Number of input reads",field),value])} %
    %
    additional samples without raw data in the \ac{SRA})
    %

  \item  \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
    different tissues

  \item Samples from Roadmap Epigenomics
    \cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping}, GTEx
    \cite{Gtexconsortium.2015:Thegenotypetissueexpression},
    uterine (GSE50599,
    \cite{Chan.ea2014:Assessmentofmyometrialtranscriptome}) and
    placenta (GSE66622,
    \cite{Hughes.ea2015:Evaluatingintra-andinter-individual})
  \item aligned to GRCh38 using STAR
    \cite{Dobin.ea2013:STARultrafastuniversalRNA-seq}
  \item called using Cufflinks 2 v2.0.10
    \cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}.

  \item Mean uniquely mapped reads: %
    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value])))}\%
  \item Mean total mapping percentage: %
      \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value]))+as.numeric(gsub("%","",star.logs[field=="% of reads mapped to multiple loci",value])))}\%.
    % 
      \item RPKMs from publicly available result files were used for
        samples where reads were not present in SRA.
   \end{itemize}

\end{frame}

\section{Genes Characteristic of Uterus}

% <<uterus.specific.markers.table,results="asis",echo=FALSE>>=
% temp <- tissue.specific.genes[tissue=="uterus",list(gene,specificity,gi.max)]
% setorder(temp,-gi.max)
% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
% print(xtable(temp,
%              caption="Uterus specific genes",label="tab:uterus_specific_genes"),
%       floating=FALSE,
%       include.rownames=FALSE,
%       )
% rm(temp)
% @ 
% }

\begin{frame}{Genes Characteristic of Uterus}
\begin{columns}
\column{0.6\textwidth}

<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("uterus")
      + xlab("Genes Characteristic of Uterus ")
      + ylab("Tissue"))
@
\column{0.4\textwidth}

\captionof{figure}{Expression of genes characteristic of uterus
  ($\tau_g \ge 0.98$) in uterus and all tissues with the second
  highest expression for any of the uterus-characterizing genes
  ordered by their tissue specificity. The known uterus-characterizing
  gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp)
  are highly characteristic of uterus.}
\end{columns}
\end{frame}

\section{Genes Characteristic of Placenta}

\begin{frame}{Genes Characteristic of Placenta}
% <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
% temp <- tissue.specific.genes[tissue=="placenta" & gi.max >= 1000,
%                               list(gene,specificity,gi.max)]
% setorder(temp,-gi.max)
% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
% print(xtable(temp,
%              caption="Placenta specific genes",label="tab:placenta_specific_genes"),
%       floating=FALSE,
%       include.rownames=FALSE,
%       )
% rm(temp)
% @

\begin{columns}
\column{0.6\textwidth}
<<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("placenta",
                            genes=as.character(tissue.specific.genes[tissue=="placenta" & 
                                                                         gi.max >= 1024
                               ,gene]))
      + xlab("Genes Characteristic of Placenta")
      + ylab("Tissue"))
@

\column{0.4\textwidth}
\captionof{figure}{Expression of genes characteristic of placenta
  ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
  in placenta and all tissues with the second highest expression for
  any placenta-characterizing gene ordered by their specificity to
  placenta. The previously known placenta-characterizing PSG genes and
  CSH genes are nearly unique to placenta and the short RNAs mir-126,
  mir-622, and PAPPA-AS1 are less so.}
\end{columns}
\end{frame}

\section{SVM to distinguish tissues}

  % Describe SVM approach

  % Figure showing the samples

  % Ability to identify placenta

  % Ability to identify uterus

\begin{frame}{Generating SVM}
  The \ac{SVM} was generated using caret
    \cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
    \cite{Karatzoglou2004:kernlab} with 10 cycles of 10 repeats and a
    tune length of 8 with a training set representing $\approx 75\%$ of the
    samples ($\approx 2450$ samples total, including $\approx 35$
    uterine or $\approx 60$ placenta) in each of the groups chosen at
    random, with the remaining samples making up a testing set
    ($\approx 820$ samples).
  \end{frame}

\begin{frame}{Placenta and Uterus SVM}
  \centering
<<placenta.roc.curve,fig.width=8,fig.height=4,out.height="0.6\\textheight",echo=FALSE>>=
layout(matrix(1:2,1,2))
pl.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"actual"])-1,
             as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"placenta"]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
pl.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"actual"])-1,
              as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"placenta"]),
              percent=TRUE,
              col="#E87722")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",
       legend=c("Placenta Testing Set", "Placenta Training Set"),
       col=c("#002058", "#E87722"),
       lwd=2)

ut.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"actual"])-1,
             as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"uterus"]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
ut.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"actual"])-1,
              as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"uterus"]),
              percent=TRUE,
              col="#E87722")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",
       legend=c("Uterus Testing Set", "Uterus Training Set"),
       col=c("#002058", "#E87722"),
       lwd=2)
@ 
\captionof{figure}{\ac{ROC} curve of the placenta \ac{SVM} (left) and
  uterus \ac{SVM} right) . Both \ac{SVM} are highly effective at
  distinguishing between samples of uterus origin and non-uterus
  origin.}
\label{fig:plac_ut_full_roc}
\end{frame}

\subsection{Identification of tissue in reduced reads}

\begin{frame}{Reducing Reads}
\end{frame}

\begin{frame}{Identification of tissue in reduced reads}
  \centering
<<placenta.roc.reduced.curve,fig.width=8,fig.height=4,out.height="0.6\\textheight",echo=FALSE>>=
layout(matrix(1:2,1,2))
for (a in c("placenta","uterus")) {
pl.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,"actual"])-1,
             as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,a]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
pl.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,a]),
              percent=TRUE,
              col="#E87722")
pl.roc.obj3 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,a]),
              percent=TRUE,
              col="green")
pl.roc.obj4 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"placenta"]),
              percent=TRUE,
              col="brown")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
       col=c("brown","green","#002058", "#E87722"),
       lwd=2)
}
@ 
\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta and
  Uterus samples with the indicated number of total reads. Reads were
  reduced by sampling reads for each sample weighted by the FPKM
  measured for each sample, and run using the same trained SVM
  depicted in Figure \ref{fig:plac_ut_full_roc}. %
  % 
  The placenta \ac{SVM} is fairly accurate at distinguishing between
  samples of placental origin and non-placental origin even when the
  number of reads is very low (10). The uterus \ac{SVM} requires many
  more reads in order to achieve similar accuracy. This is likely due
  to the much smaller number of uterus specific genes in comparison to
  placenta.}

\end{frame}

\section{Conclusions}

\begin{frame}{Conclusions}
  \begin{itemize}
  \item Pregnancy-specific glycoproteins (PSG1, 2, 3 et al.) are
    characteristic of placental tissues, and PAEP is characteristic of
    uterus, as previously described.
%   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
  \item \ac{SVM} are able to accurately identify the tissue of origin
    for uterus and placenta.
  \item Very few sequenced reads are required to accurately identify
    tissues of Placental origin. We are able to detect 100\% of the
    placenta tissues with 80\% specificity with just 10 reads.
  \item We can easily identify exosomes originating from the placenta,
    as even the smallest exosomes can contain 70 RNA molecules
    \cite{Li.ea2014:AnalysisofRNAcontent}.
  \item Non-maternal variants may also be able to distinguish between
    placental reads and maternal reads.
  \end{itemize}
\end{frame}

\section*{Acknowledgements}

\begin{frame}
  The code and underlying data for this poster can be found at
\begin{center}
\qrcode[padding]{http://dla2.us/p/em2015}
\url{http://dla2.us/p/em2015}
\end{center}
\end{frame}

\section*{References}

\begin{frame}[plain]{References}
  \begin{center}
    \mbox{}\vspace{-\baselineskip}
    \printbibliography[heading=none]
  \end{center}
\end{frame}

\begin{frame}{Acronyms}
  \acrodef{EV}[EV]{Extracellular Vesicle}
\acrodef{SVM}{Support Vector Machine}
\acrodef{ROC}{Receiver Operating Characteristic}
\acrodef{KNN}{K Nearest Neighbors}
\acrodef{SRA}{Sequence Read Archive}

\end{frame}

\end{document}