### 2015-10-12 slides/slides.Rnw
add better slides for ISRS presentation; fix date

\documentclass[ignorenonframetext]{beamer}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{booktabs}
% \usepackage{multirow}
\usepackage{acronym}
\usepackage{qrcode}
\usepackage[absolute,overlay]{textpos}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
\addbibresource{references.bib}
% \usepackage[nomargin,inline,draft]{fixme}
% \newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
% \usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
\usepackage{texshade}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\renewcommand*{\bibfont}{\tiny}
\usepackage{mathtools,cancel}

\renewcommand{\CancelColor}{\color{red}} %change cancel color to red

\makeatletter
\let\my@cancelto\cancelto %copy over the original cancelto command
\newcommand<>{\cancelto}[2]{\alt#3{\my@cancelto{#1}{#2}}{\mathrlap{#2}\phantom{\my@cancelto{#1}{#2}}}}
% redefine the cancelto command, using \phantom to assure that the
% result doesn't wiggle up and down with and without the arrow
\makeatother

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.  Use showboxes option to show outlines of textboxes.
% \usepackage[absolute]{textpos}
\usepackage[absolute,showboxes]{textpos}

\usepackage{multirow}
\usepackage{array}

\mode<presentation>{ 
  \usetheme{CambridgeUS}
  \usecolortheme{crane}
%   % http://identitystandards.illinois.edu/graphicstandardsmanual/generalguidelines/colors.html
  \definecolor{ilboldblue}{HTML}{002058}
  \definecolor{ilboldorange}{HTML}{E87722}
  \definecolor{ilblue}{HTML}{606EB2}
  \definecolor{ilorange}{HTML}{D45D00}
%   \setbeamercolor{alerted text}{fg=ilboldorange}
%   \setbeamercolor*{palette primary}{fg=ilboldblue,bg=ilorange!70!white}
%   \setbeamercolor*{palette secondary}{fg=ilboldblue,bg=ilorange!60!white}
%   \setbeamercolor*{palette tertiary}{bg=ilblue!60!white,fg=ilboldorange}
%   \setbeamercolor*{palette quaternary}{fg=ilboldblue,bg=ilorange!80!white}
%   
%   \setbeamercolor*{sidebar}{fg=ilorange,bg=ilboldblue}
%   
%   \setbeamercolor*{palette sidebar primary}{fg=ilblue!10!white,bg=ilorange}
%   \setbeamercolor*{palette sidebar secondary}{fg=ilorange}
%   \setbeamercolor*{palette sidebar tertiary}{fg=ilblue}
%   \setbeamercolor*{palette sidebar quaternary}{fg=ilorange}
%   
%   % \setbeamercolor*{titlelike}{parent=palette primary}
%   \setbeamercolor{titlelike}{parent=palette primary,fg=ilboldblue,bg=ilorange}
%   \setbeamercolor{frametitle}{fg=ilboldorange,bg=ilblue!20!white}
%   \setbeamercolor{frametitle right}{fg=ilboldblue,bg=ilorange}
%   
%   \setbeamercolor*{separation line}{}
%   \setbeamercolor*{fine separation line}{}
%   \setbeamercovered{transparent}  
  \logo{\begin{tikzpicture}% Pale figure
      {\node[opacity=0.734]{\IfFileExists{figures/igb_wordmark.pdf}{\includegraphics[width=2cm,height=1cm,keepaspectratio]{figures/igb_wordmark}}{}%
        };}%
    \end{tikzpicture}}
}

\title[EVs of Placenta and Uterus]{Identifying Extracellular Vesicles of Placenta and Uterine 
  Origin Using RNA Expression Signatures}
\author[Don Armstrong et. al.]{Don L. Armstrong, Monica Uddin, Derek Wildman}
\institute[IGB]{Institute for Genomic Biology, Computing Genomes 
  for Reproductive Health, University of Illinois, Urbana-Champaign}
\subject{RNA expression}
\date{October 12, 2015}
\begin{document}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.8\\textheight",out.extra="keepaspectratio")
#opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(digits=2)
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
library("grid")
library("xtable")
library("caret")
library("pROC")
library("sampling")

capwords <- function(s, strict = FALSE) {
    cap <- function(s) paste(toupper(substring(s, 1, 1)),
                             {s <- substring(s, 2); if(strict) tolower(s) else s},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}
@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced")))>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
@

\IfFileExists{figures/relevant_xkcd.png}{\frame[plain]{\centering \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{figures/relevant_xkcd.png}

    \url{https://xkcd.com/1217/}}}

\frame[plain]{\titlepage}

<<specific.markers.plot.function,echo=FALSE,results="hide",warning=FALSE>>=
specific.markers.plot <- function(tissue.name,genes) {
    setkey(interesting.gene.reads.wide,"gene_short_name")
    if (missing(genes))
        genes <- as.character(tissue.specific.genes[tissue==tissue.name,gene])
    interesting.tissues <- 
        colnames(interesting.gene.reads.wide)[
         apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
                             function(x){order(-x)[2]+2})]

    interesting.tissues <- 
        c("gene_short_name",tissue.name,unique(interesting.tissues))

    tissue.specific.gene.reads <- 
        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
                        id="gene_short_name",
                        variable.name="tissue",
                        value.name="FPKM"
                        ))
    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
    gene.levels <- 
        unique(as.character(tissue.specific.genes[tissue==tissue.name][order(specificity),gene]))
    tissue.specific.gene.reads[,gene_short_name:=factor(gene_short_name,levels=gene.levels)]
    return(ggplot(tissue.specific.gene.reads,
                  aes(y=gene_short_name,
                          x=tissue))
           + geom_tile(aes(fill=`log_2(FPKM)`))
           + scale_fill_gradient(limits=c(0,17),
                                 low="white",
                                 high="#002058",
                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
           + scale_x_discrete("", expand = c(0, 0))
           + scale_y_discrete("", expand = c(0, 0))
           + theme_grey(base_size=14)
           + theme(legend.position = "top",
                   axis.ticks = element_blank(),
                   panel.grid=element_blank(),
                   panel.border=element_blank(),
                   panel.background=element_blank(),
                   plot.background=element_blank(),
                   plot.margin=unit(c(1,1,1,1),"mm"),
                   axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
           )
}
@

\section{Introduction}

\begin{frame}{Introduction: \acsp{EV}}
  \acp{EV} are
  \begin{itemize}
  \item small (40-1000nm) compartments
  \item found in blood, saliva, and urine. 
  \item originate from the plasma membrane or multivesicular endosome
  \item from many cell types
  \item contain proteins, lipids, mRNA, miRNA, and other ncRNAs.
  \item \alert{remote sensor of the transcriptome of inaccessible tissues}
  \end{itemize}
\end{frame}

\subsection{Identifying source tissue of \acp{EV}}

\begin{frame}{Identifying source tissue of \acp{EV}}
  \begin{enumerate}
  \item We want to remotely sense the transcriptome of tissues
  \item For each \ac{EV}, identify the source tissue

  \item So, identify miRNA and mRNA markers which are characteristic of
    multiple organs and tissues
    \begin{itemize}    
    \item Use publicly-available RNAseq transcriptomes

    \item Identify characteristic markers by calculating a tissue
      specificity index for each gene $g$ ($\tau_g$):

      \begin{equation}
        \tau_g = \frac{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}{N-1}
      \end{equation}
      \label{eq:tissue_specificity_index}
 \item $N$ is the number of tissues, $X_{i,g}$ is the mean expression
   of gene $g$ in tissue $i$ and $\mathrm{max}(X_g)$ is the maximal
   expression of gene $g$
    \end{itemize}
\end{enumerate}
\only<2-4>{
\begin{textblock*}{64mm}(32mm,0.25\textheight)
\begin{block}{Tissue Specificity}
  \begin{itemize}
    \item Highly specific $\tau_g = \frac{\cancelto<3->{N-1}{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}}{N-1} = 1$
    \item Not specific $\tau_g = \frac{\cancelto<4->{0}{\sum_{i=1}^{i=N} \left(1-\frac{X_{i,g}}{\mathrm{max}(X_g)}\right)}}{N-1} = 0$
    \end{itemize}
\end{block}
\end{textblock*}
}
\end{frame}

\begin{frame}{Source of \acp{EV} --- Machine Learning}
  \begin{itemize}
  \item High tissue specificity might not be the best discriminator
  \item Use machine learning techniques!
  \item Train a \ac{SVM}
  \item Use genes which are highly expressed and have either
    \begin{itemize}
    \item High Tissue Specific Index ($\tau_g$)
    \item High Entropy
    \end{itemize}
  \end{itemize}
\end{frame}

\section{Methods}
\begin{frame}{Methods}
  \begin{center}
  \begin{tikzpicture}
    [rawreads/.style={rectangle,rounded corners=2mm,draw=ilblue!90!white,fill=ilblue!40!white,inner sep=2pt,minimum size=8mm,thick},
    processing/.style={rectangle,rounded corners=2mm,draw=ilblue!90!white,fill=ilboldblue!40!white,inner sep=2pt,minimum size=8mm,thick},
    arrow/.style={->,shorten >=1pt,>=stealth,thick},
    tissues/.style={rectangle,rounded corners=2mm,draw=ilorange!90!white,fill=ilorange!40!white,inner sep=2pt,minimum size=8mm,thick}]
    \node[tissues] (tissues) {\Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} tissues};
    \node[processing] (star) [above left=of tissues] {STAR+Cufflinks};
    \node[processing] (fpkm) [above right=of tissues] {FPKMs published};
    \node[rawreads] (samples) [above=of star] {\Sexpr{length(star.logs[grepl("Number of input reads",field),value])} samples in SRA};
    \node[rawreads] (samples noraw) [above=of fpkm] {\Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
        nrow(predictions.caret.svm[["uterus"]][["testing"]]) -
        length(star.logs[grepl("Number of input reads",field),value])} samples not in SRA};
    \node[processing] (matrix) [below=of tissues] {FPKMs of 63967 genes $\times$ \Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
        nrow(predictions.caret.svm[["uterus"]][["testing"]])} samples};
    \draw[arrow] (samples)--(star);
    \draw[arrow] (star)--(tissues);
    \draw[arrow] (samples noraw)--(fpkm);
    \draw[arrow] (fpkm)--(tissues);
    \draw[arrow] (tissues)--(matrix);
  \end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}<beamer:0>{Methods}
  \begin{itemize}
  % 
  \item \Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]} %
    %
    RNAseq reads (\Sexpr{star.logs[grepl("Number of input reads",field),sum(as.numeric(value))]/length(star.logs[grepl("Number of input reads",field),value])} average per sample)%
    % 
  \item  \Sexpr{length(star.logs[grepl("Number of input reads",field),value])} %
    % 
    samples (supplemented with FPKMs from
    \Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
      nrow(predictions.caret.svm[["uterus"]][["testing"]]) -
      length(star.logs[grepl("Number of input reads",field),value])} %
    %
    additional samples without raw data in the \ac{SRA})
    %

  \item  \Sexpr{length(unique(na.omit(tissue.specific.genes[,tissue])))} %
    different tissues

  \item Samples from Roadmap Epigenomics
    \cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping}, GTEx
    \cite{Gtexconsortium.2015:Thegenotypetissueexpression},
    uterine (GSE50599,
    \cite{Chan.ea2014:Assessmentofmyometrialtranscriptome}) and
    placenta (GSE66622,
    \cite{Hughes.ea2015:Evaluatingintra-andinter-individual})
  \item aligned to GRCh38 using STAR
    \cite{Dobin.ea2013:STARultrafastuniversalRNA-seq}
  \item quantified using Cufflinks 2 v2.0.10
    \cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions}.

  \item Mean uniquely mapped reads: %
    \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value])))}\%
  \item Mean total mapping percentage: %
      \Sexpr{mean(as.numeric(gsub("%","",star.logs[field=="Uniquely mapped reads %",value]))+as.numeric(gsub("%","",star.logs[field=="% of reads mapped to multiple loci",value])))}\%.
    % 
      \item RPKMs from publicly available result files were used for
        samples where reads were not present in SRA.
   \end{itemize}

\end{frame}

\section{Genes Characteristic of Uterus}

% <<uterus.specific.markers.table,results="asis",echo=FALSE>>=
% temp <- tissue.specific.genes[tissue=="uterus",list(gene,specificity,gi.max)]
% setorder(temp,-gi.max)
% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
% print(xtable(temp,
%              caption="Uterus specific genes",label="tab:uterus_specific_genes"),
%       floating=FALSE,
%       include.rownames=FALSE,
%       )
% rm(temp)
% @ 
% }

\begin{frame}{Genes Characteristic of Uterus}
\begin{columns}
\column{0.7\textwidth}

<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("uterus")
      + xlab("Genes Characteristic of Uterus ")
      + ylab("Tissue"))
@
\column{0.3\textwidth}

\captionof{figure}{Expression of genes characteristic of uterus
  ($\tau_g \ge 0.98$) in uterus and tissues with second-highest
  expression. The known uterus-characterizing gene
  progestagen-associated endometrial protein (PAEP) and short RNAs
  ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp) are highly characteristic
  of uterus.}
\end{columns}
\end{frame}

\section{Genes Characteristic of Placenta}

\begin{frame}{Genes Characteristic of Placenta}
% <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
% temp <- tissue.specific.genes[tissue=="placenta" & gi.max >= 1000,
%                               list(gene,specificity,gi.max)]
% setorder(temp,-gi.max)
% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
% print(xtable(temp,
%              caption="Placenta specific genes",label="tab:placenta_specific_genes"),
%       floating=FALSE,
%       include.rownames=FALSE,
%       )
% rm(temp)
% @

\begin{columns}
\column{0.6\textwidth}
<<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("placenta",
                            genes=as.character(tissue.specific.genes[tissue=="placenta" & 
                                                                         gi.max >= 1024
                               ,gene]))
      + xlab("Genes Characteristic of Placenta")
      + ylab("Tissue"))
@

\column{0.4\textwidth}
\captionof{figure}{Expression of genes characteristic of placenta
  ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
  in placenta and tissues with the second highest expression. The
  previously known placenta-characterizing PSG genes and CSH genes are
  nearly unique to placenta and the short RNAs mir-126, mir-622, and
  PAPPA-AS1 are less so.}
\end{columns}
\end{frame}

\section{SVM to distinguish tissues}

  % Describe SVM approach

  % Figure showing the samples

  % Ability to identify placenta

  % Ability to identify uterus

\subsection{SVM Methodology}

\begin{frame}{Generating SVM}
 \begin{itemize}
 \item \ac{SVM} is a machine learning technique which finds the
   hyperplane with the maximal margins which separates two groups
 \item \ac{SVM} was generated using caret
    \cite{Kuhn2008:Buildingpredictivemodelsinr} and svmRadial
    \cite{Karatzoglou2004:kernlab}
  \item 10 cycles of 10 repeats and a
    tune length of 8
  \end{itemize}
\end{frame}

\begin{frame}{Testing and Training Sets}
  \begin{center}
  \begin{tikzpicture}
    [training/.style={rectangle,rounded corners=2mm,draw=ilblue!90!white,fill=ilblue!40!white,inner sep=2pt,minimum size=8mm,thick},
    testing/.style={rectangle,rounded corners=2mm,draw=ilblue!90!white,fill=ilboldblue!40!white,inner sep=2pt,minimum size=8mm,thick},
    arrow/.style={->,shorten >=1pt,>=stealth,thick},
    sample/.style={rectangle,rounded corners=2mm,draw=ilorange!90!white,fill=ilorange!40!white,inner sep=2pt,minimum size=8mm,thick}]
    \node[sample] (samples) {\Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])+
        nrow(predictions.caret.svm[["uterus"]][["testing"]])} samples};
\onslide<2->{    \node[training] (training) [below left=of samples] {\Sexpr{nrow(predictions.caret.svm[["uterus"]][["training"]])} training samples};
    \node[training] (training uterus) [below=of training] {\Sexpr{sum(predictions.caret.svm[["uterus"]][["training"]][,"actual"]=="uterus")} uterus,
      \Sexpr{sum(predictions.caret.svm[["placenta"]][["training"]][,"actual"]=="placenta")} placenta};

    \draw[arrow] (samples) -- node[above left=8pt]{75\%} (training);
    \draw[arrow] (training) -- (training uterus);
    }
\onslide<3->{    \node[testing] (testing) [below right=of samples] {\Sexpr{nrow(predictions.caret.svm[["uterus"]][["testing"]])} testing samples};
    \node[testing] (testing uterus) [below=of testing] {\Sexpr{sum(predictions.caret.svm[["uterus"]][["testing"]][,"actual"]=="uterus")} uterus,
      \Sexpr{sum(predictions.caret.svm[["placenta"]][["testing"]][,"actual"]=="placenta")} placenta};
    \draw[arrow] (samples) -- node[above right=8pt]{25\%} (testing);
    \draw[arrow] (testing) -- (testing uterus);}
  \end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}{Placenta SVM: Training and Testing}
  \begin{columns}
    \column{0.6\textwidth}
<<placenta.roc.curve,fig.width=4,fig.height=4,out.height="0.7\\textheight",echo=FALSE>>=
# layout(matrix(1:2,1,2))
par(mar=c(0,0,0,0))
pl.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"actual"])-1,
             as.numeric(predictions.caret.svm[["placenta"]][["testing"]][,"placenta"]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
pl.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"actual"])-1,
              as.numeric(predictions.caret.svm[["placenta"]][["training"]][,"placenta"]),
              percent=TRUE,
              col="#E87722")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",
       legend=c("Placenta Testing Set", "Placenta Training Set"),
       col=c("#002058", "#E87722"),
       lwd=2)
@ 
\column{0.4\textwidth}
\captionof{figure}{\ac{ROC} curve of the placenta \ac{SVM}. The
  \ac{SVM} is highly effective at distinguishing between samples of
  placental origin and non-placental origin.}
\label{fig:plac_full_roc}
\end{columns}
\end{frame}

\begin{frame}{Uterus SVM: Training and Testing Sets}
  \begin{columns}
    \column{0.6\textwidth}
<<uterus.roc.curve,fig.width=4,fig.height=4,out.height="0.7\\textheight",echo=FALSE>>=

ut.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"actual"])-1,
             as.numeric(predictions.caret.svm[["uterus"]][["testing"]][,"uterus"]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
ut.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"actual"])-1,
              as.numeric(predictions.caret.svm[["uterus"]][["training"]][,"uterus"]),
              percent=TRUE,
              col="#E87722")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",
       legend=c("Uterus Testing Set", "Uterus Training Set"),
       col=c("#002058", "#E87722"),
       lwd=2)
@ 
\column{0.4\textwidth}
\captionof{figure}{\ac{ROC} curve of the uterus \ac{SVM}. The \ac{SVM}
  is highly effective at distinguishing between samples of uterus
  origin and non-uterus origin.}
\label{fig:ut_full_roc}
\end{columns}
\end{frame}

\subsection{Identification of tissue in reduced reads}

\begin{frame}{Reducing Reads}
  \begin{itemize}
  \item Reads sub-sampled from each sample 
  \item Weighted by the FPKM measured
  \item Run using the same trained SVM shown previously
  \end{itemize}
  \only<2->{
\begin{textblock*}{100mm}(20mm,0.25\textheight)
\begin{block}{Sub sampling}
<<calc.reduced,echo=FALSE,results="asis">>=
reduced.reads <- function(fpkms,num.reads) {
    non.zero <- fpkms > 0
    ret.val <- rep.int(0,length(fpkms))
    if (!any(non.zero)) {
        return(fpkms)
    }
    inc.prob <- inclusionprobabilities(fpkms[non.zero],1)
    ret.val[non.zero] <-
        apply(sapply(1:(num.reads/1),function(y){UPrandomsystematic(inc.prob,0)}),1,sum)
    names(ret.val) <- names(fpkms)
    ret.val
}
set.seed(1021)
reads <- c("A"=1000,
           "B"=2000,
           "C"=20,
           "D"=5000)
reads.to.table <- function(read,reads){
    data.table("Total Reads"=read,
               "Gene A"=reads["A"],
               "Gene B"=reads["B"],
               "Gene C"=reads["C"],
               "Gene D"=reads["D"])
}
rr.table <- list(reads.to.table("Orig. FPKM",reads))
i <- 0
for (read in c(500,500,100,100,5,5,5)){
    i <- i+1
    reads <- reduced.reads(reads,read)
    rr.table[[paste0("read.",read,'.',i)]] <- 
        reads.to.table(read,reads)
}
print(xtable(rbindlist(rr.table),
             digits=0
             ),
      include.rownames=FALSE)
@

\end{block}
\end{textblock*}
}

\end{frame}

\begin{frame}{Identification of tissue in reduced reads}

\begin{columns}
\column{0.6\textwidth}
<<placenta.roc.reduced.curve,fig.width=6,fig.height=6,out.height="0.8\\textheight",echo=FALSE>>=
#layout(matrix(1:2,1,2))
for (a in c("placenta")) {
pl.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,"actual"])-1,
             as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,a]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
pl.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,a]),
              percent=TRUE,
              col="#E87722")
pl.roc.obj3 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,a]),
              percent=TRUE,
              col="green")
pl.roc.obj4 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"placenta"]),
              percent=TRUE,
              col="brown")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
       col=c("brown","green","#002058", "#E87722"),
       lwd=2)
}
@ 
\column{0.4\textwidth}
\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta with the indicated number of total reads.  %
  % 
  The placenta \ac{SVM} is fairly accurate at distinguishing between
  samples of placental origin and non-placental origin even when the
  number of reads is very low (10).}
\end{columns}
\end{frame}

\begin{frame}{Reduced Reads: Uterus}
\begin{columns}
\column{0.6\textwidth}
<<uterus.roc.reduced.curve,fig.width=6,fig.height=6,out.height="0.8\\textheight",echo=FALSE>>=
# layout(matrix(1:2,1,2))
for (a in c("uterus")) {
pl.roc.obj1 <- 
    plot.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,"actual"])-1,
             as.numeric(predictions.caret.svm.reduced[[a]][["testing.1000.reads"]][,a]),
             percent=TRUE,
             ## main="Training vs Testing set for Uterus",
             col="#002058"
             )
pl.roc.obj2 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[[a]][["testing.5000.reads"]][,a]),
              percent=TRUE,
              col="#E87722")
pl.roc.obj3 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[[a]][["testing.10.reads"]][,a]),
              percent=TRUE,
              col="green")
pl.roc.obj4 <- 
    lines.roc(as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"actual"])-1,
              as.numeric(predictions.caret.svm.reduced[["placenta"]][["testing.1.reads"]][,"placenta"]),
              percent=TRUE,
              col="brown")
# testobj <- roc.test(rocobj1, rocobj2)
# text(50, 50, labels=paste("p-value =", format.pval(testobj$p.value)), adj=c(0, .5))

legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
       col=c("brown","green","#002058", "#E87722"),
       lwd=2)
}
@ 
\column{0.4\textwidth}
\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta and
  Uterus samples with the indicated number of total reads.  %
  % 
  The uterus \ac{SVM} requires many more reads in order to achieve
  similar accuracy. This is likely due to the much smaller number of
  uterus specific genes in comparison to placenta.}
\end{columns}
\end{frame}

\section{Conclusions}

\begin{frame}{Conclusions}
  \begin{itemize}
  \item Pregnancy-specific glycoproteins (PSG1, 2, 3 et al.) are
    characteristic of placental tissues, and PAEP is characteristic of
    uterus, as previously described.
%   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
  \item \ac{SVM} are able to accurately identify the tissue of origin
    for uterus and placenta.
  \item Very few sequenced reads are required to accurately identify
    tissues of Placental origin. We are able to detect 100\% of the
    placenta tissues with 80\% specificity with just 10 reads.
  \item We can easily identify exosomes originating from the placenta,
    as even the smallest exosomes can contain 70 RNA molecules
    \cite{Li.ea2014:AnalysisofRNAcontent}.
  \item Non-maternal variants may also be able to distinguish between
    placental reads and maternal reads.
  \end{itemize}
\end{frame}

\section*{Acknowledgments}

\begin{frame}{Acknowledgments}
  \begin{itemize}
  \item Derek Wildman
  \item Monica Uddin
  \end{itemize}

  The code and underlying data for this talk can be found at
  \begin{center}
    \qrcode[padding]{http://dla2.us/p/em2015}
    \url{http://dla2.us/p/em2015}
  \end{center}
  \begin{itemize}
  \item University of Illinois and Institute for Genomic Biology
    Startup Funds to Derek and Monica
  \end{itemize}
\end{frame}

\begin{frame}[plain]
  \acrodef{EV}[EV]{Extracellular Vesicle}
  \acrodef{SVM}{Support Vector Machine}
  \acrodef{ROC}{Receiver Operating Characteristic}
  \acrodef{KNN}{K Nearest Neighbors}
  \acrodef{SRA}{Sequence Read Archive}

\end{frame}

\section*{References}

\begin{frame}[plain]{References}
  \begin{center}
    \mbox{}\vspace{-\baselineskip}
    \printbibliography[heading=none]
  \end{center}
\end{frame}

\end{document}